### 2018-01-29 cabios-template/main.tex
start drafting manuscript

\documentclass{bioinfo}
\copyrightyear{2015} \pubyear{2015}

%https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3463124/

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Subject Section}

\title[Rethomics]{Rethomics: an R framework to analyse high-throughput behavioural data}
\author[Geissmann \textit{et~al}.]{Quentin Geissmann\,$^{\text{\sfb 1,}*}$, Luis Garcia\,$^{\text{\sfb 2}}$, Esteban Beckwith\,$^{\text{\sfb 1}}$,  and Giorgio Gilestro\,$^{\text{\sfb 1,}*}$}
\address{$^{\text{\sf 1}}$Department, Institution, City, Post Code, Country and \\
$^{\text{\sf 2}}$Department, Institution, City, Post Code,
Country.}

\corresp{$^\ast$To whom correspondence should be addressed.}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\abstract{
	\textbf{Motivation:}
		Ethomics, a quantitative and high-throughput approach to ethology, is a novel and exciting field.
		The recent development of methods that automatically score variables in multiple animals provides an unprecedented insight into the study of behaviours. 
		The analysis of ethomics data presents many challenges that are conceptually independent of the acquisition platform.
		However, there is, little effort in providing a general framework to analyse multiple and long behavioural time series.
		We developped the  Rethomics framework, a suite of R packages that altogether provide utilities to:
		import, store, visualise and analyse behavioural data.\\
	\textbf{Availability and Implementation:}
	All packages in the rethomics framework are available under the terms of the GPLv3 license.
	Exaustive installation instructions and tutorials are available at
 		\href{https://github.com/rethomics}{https://github.com/rethomics/} 
 		\href{https://rethomics.github.io/}{https://rethomics.github.io/}.\\
	\textbf{Contact:}
		\href{qgeissmann@gmail.com}{qgeissmann@gmail.com}
	}

\maketitle

\section{Introduction}
TODOs:

	1. Ethomics, a scaling up in time but also in number of animals.
	2. A diversity of platform to do the same thing.
	3. Need for well documented programmatic interface  innovative/advances analysis
	4. The idea of unifying behavioural analysis, so community (experimental and quantitative biologists) collaborates

%\enlargethispage{12pt}

\section{Approach}

Equation~(\ref{eq:01}) Text Text Text Text Text Text  Text Text
Text Text Text Text Text Text Text Text Text Text Text Text Text.
Figure~2\vphantom{\ref{fig:02}} shows that the above method  Text
Text Text Text  Text Text Text Text Text Text  Text Text.
\citealp{Boffelli03} might want to know about text text text text

\section{Framework  features}
\subsection{Data import}
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
\subsection{Internal data structure}
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
\subsection{Circadian analysis}
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
\subsection{Sleep scoring}
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
\subsection{Visualisation}
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl

\section{Conclusion}

xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl 
xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl xxx xxxxx xxxx dwjdkl

\section*{Acknowledgements}

Han Kim. Patrick Kr{\"a}tschmer

\section*{Funding}

This work has been supported by the... Text Text  Text Text.\vspace*{-12pt}

%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}

\begin{thebibliography}{}

\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.

\bibitem[Bag {\it et~al}., 2001]{Bag01}
Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.

\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
in nigral dopaminergic neurnol cell: correlation with the known
pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
Res.}, \textbf{110}(Suppl. 1), 76--84.

\bibitem[Lehmann, 1986]{Leh86}
Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.

\bibitem[Crenshaw and Jones, 2003]{Cre03}
Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
cancer management: one tumor, one chip. \textit{Bioinformatics},
doi:10.1093/bioinformatics/btn000.

\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
???--???.

\bibitem[Bardet, 1920]{Bar20}
Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
polydactylie et retinite pigmentaire (contribution a l'etude des
formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
institution, Paris, France.

\end{thebibliography}
\end{document}