### 2018-02-08 plos-latex-template/results/script.Rmd
result script

---
output:
  pdf_document:
    keep_tex: true
---

# Results

## Description
In order to illustrate some mechanisms in rethomics, we provide a simple and reproducible example of analysis of 
circadian phenotype recorded with DAM2 monitors -- a widely addopted paradigm.

TODO:
* data source -- The data was obtained from ...(citation) TODO.
* description (genotypes etc) -- exhaustive description in the `metadata.csv`

## Data loading
First of all, the necessary `rethomics` packages are loaded.

```{r, message=FALSE}
library(damr)      # input DAM2 data
library(zeitgebr)  # periodogram computation
library(sleepr)    # sleep analysis
library(ggetho)    # behaviour visualisation
```

Then, the metadata file is read and linked to the `.txt` result files.
```{r}
met <- damr::link_dam2_metadata("./metadata.csv", ".") # linking
dt <- load_dam2(met)                                   # loading
summary(dt)                                            # quick summary
```

## Preprocessing

We notice, from the metadata, that the two replicates do not have the same time in baseline.
We would like to express the time relative to the important event: thetransition to `LL`. 
The best way is to sustract the `baseline_days` metavariable from the `t` variable.
This gives us an opportunity to illustrate the use `xmv()` that maps metavariables as variables.
In addition, we use the `data.table` syntax to create, in place, a `moving` variable.
It is `TRUE` when and only when `activity` is greater than zero:
```{r}
dt[,t := t - days(xmv(baseline_days))]    # baseline sustraction. not the use of xmv
dt[, moving :=  activity > 0]  
```

To simplify visualisation, we create our own `label` metavariable, as combination of a number and `genotype`.
In the restricted context of this analysis, this acts a unique identifyier.
Importanly, we keep `id` , which is more rigourous and universal.

```{r}
dt[, label := interaction(1:.N, genotype), meta=T]
# print(dt)
```

## Curation

It is important to see an overview of how each individual and experiment behaved and, if necessary, alter the data accordingly. We save this figure as Fig 3A.

```{r}
# make a ggplot object with label on the y and moving on the z axis
fig3A <- ggetho(dt, aes(y=label, z=moving)) +  
  # show data as a tile plot. That is z is a pixel whose intensity maps moving
  stat_tile_etho() +                  
  # add layers to draw annotations to show L and D phases as white and black
  # the first layer is for the baseline (until t=0)
  stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
  # in the 2nd one, we start at 0 and use grey instead of black as we work in LL
  stat_ld_annotations(x_limits = c(0, dt[,max(t)]), ld_colours = c("white", "grey"))
```

Dead or escaped animals are falsely scored as long series of zero-activity.
Our `sleepr` packges offer a tool to detect and remove this artifactual data:

```{r}
dt <- sleepr::curate_dead_animals(dt, moving)
# make a ggplot object with label on the y and moving on the z axis
fig3B <- ggetho(dt, aes(y=label, z=moving)) +  
    stat_tile_etho() +                  
    stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
    stat_ld_annotations(x_limits = c(0, dt[,max(t)]), ld_colours = c("white", "grey"))
```
The updated version can be visualised in Fig 3B.

For the purpose of this example, we keep only individuals that have at least five days in LL.
```{r}
valid_dt <- dt[ , .(valid = max(t) > days(5)), by=id]
valid_ids <- valid_dt[valid == T, id]
dt <- dt[id %in% valid_ids]
summary(dt)
```
Note that as a result, we now have `r nrow(dt[meta=T])` "valid" individuals.

## Double plotted actograms

A common way of representing rythmicity in circadian experiments is to compute "double-plotted actograms".
In Fig S1A, we show all double plotted actograms layed out in a grid.

```{r}
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            stat_bar_tile_etho() + 
            facet_wrap( ~ label, ncol=4) +
            scale_y_discrete(name="Day")
```

## Periodograms

For each indidual, we compute a $\chi{}^2$ periodogram and we layout all of them in a grid (Fig S1B).

```{r}
dt_ll <- dt[t > days(1)]
per_dt <- periodogram(moving, 
						dt_ll, 
						resample_rate = 1/mins(10),
						FUN=chi_sq_periodogram)

per_dt <- find_peaks(per_dt)

figS1B <- ggperio(per_dt, aes(y = power, peak=peak)) + 
      geom_line() +
      geom_line(aes(y=signif_threshold), colour="red") + 
      geom_peak() + 
      facet_wrap( ~ label, ncol=4)

```

```{r, eval=TRUE, echo=FALSE, message=FALSE}
################################ CODE BELOW FOR COSMETIC ALTERATIONS OF FIGURES

library(cowplot)

my_theme <- theme(axis.title.y=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.line = element_blank(),
            panel.border = element_blank(),
            legend.position="none")

legend <- fig3A + theme(axis.title=element_blank(),
                        axis.text.x=element_blank(),
                        axis.ticks=element_blank(),
                        axis.line = element_blank(),
                        #axis.ticks.y=element_blank(),
                        panel.border = element_blank(),
                        #legend.text = "moving",
                        legend.position="left") + 
                  coord_cartesian(xlim=-days(c(10, 9))) + scale_y_discrete(position = "right")

fig3 <- cowplot::plot_grid(
                          legend,
                          fig3A +  my_theme,
                          fig3B +  my_theme,
                          nrow=1,
                          labels=c("", "A", "B"),
                          rel_widths = c(1, 2, 2),
                          label_size = 20,
                          axis="b",
                          align="h"
                          )
pdf("fig3.pdf",w=12,h=12)
print(fig3)
dev.off()

figS1A <- figS1A +  theme(strip.text.x = element_text(size =8))
figS1B <- figS1B +  theme(strip.text.x = element_text(size =8))

figS1 <- cowplot::plot_grid(
                          figS1A,
                          figS1B,
                          nrow=1,
                          labels=c( "A", "B"),
                          label_size = 20,
                          align="v"
                          )

pdf("figS1.pdf",w=16,h=16)
print(figS1)
dev.off()
```