### 2018-02-08 plos-latex-template/results.Rnw
another push

<<echo=FALSE, message=FALSE>>=
opts_knit$set(root.dir = "results")
@

\section*{Results}

<<message=FALSE>>=
library(damr)      # input DAM2 data
library(zeitgebr)  # periodogram computation
library(sleepr)    # sleep analysis
library(ggetho)    # behaviour visualisation
@

Then, the metadata file is read and linked to the \texttt{.txt} result files.

<<>>=
metadata <- link_dam2_metadata("metadata.csv",".")    # linking
# print(metadata)                                     # check metadata
dt <- load_dam2(metadata)                             # loading
summary(dt)                                           # quick summary
@

\subsection*{Preprocessing}
We notice, from the metadata, that the two replicates do not have the same time in baseline.
We would like to express the time relative to the important event: the transition to \texttt{LL}. 
To do so, we subtract the \texttt{baseline\_days} metavariable from the \texttt{t} variable.
This gives us an opportunity to illustrate the use \texttt{xmv()}, which expands metavariables as variables.
In addition, we use the \texttt{data.table} syntax to create, in place, a \texttt{moving} variable.
It is \texttt{TRUE} when and only when \texttt{activity} is greater than zero:

<<message=F, results="hide">>=
# baseline subtraction -- note the use of xmv
dt[,t := t - days(xmv(baseline_days))]
dt[, moving :=  activity > 0]  
summary(dt)
@

To simplify visualisation, we create our own \texttt{label} metavariable, as combination of a number and \texttt{genotype}.
In the restricted context of this analysis, \texttt{label} acts a unique identifier.
Importantly, we keep \texttt{id} , which is more rigorous and universal.

<<message=FALSE, rsults="hide">>=
dt[, label := interaction(1:.N, genotype), meta=T]
# print(dt)
@

\subsection*{Curation}
It is important to visualise an overview of how each individual behaved and, if necessary, alter the data accordingly. For this, we generate a tile plot (Fig~\ref{fig:fig-3}A).

\begin{figure}[!h]
	\includegraphics[width=1\textwidth]{fig/fig-3.pdf}
	\caption{{\bf Experiment quality control.}
			Tile plot representing the fraction of time spent moving as a colour intensity.
			Each individual is represented by a row and time, on the x axis, is binned in 30 minute.}
	\label{fig:fig-3}
\end{figure}

<<>>=
# make a ggplot object with label on the y and moving on the z axis
fig3A <- ggetho(dt, aes(y=label, z=moving)) +  
  # show data as a tile plot
  # that is z is a pixel whose intensity maps moving
  stat_tile_etho() +                  
  # add layers to draw annotations to show L and D phases
  # as white and black, respectivelly
  # the first layer is for the baseline (until t=0)
  stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
  # in the 2nd one, we start at 0 and use grey 
  # instead of black as we work in LL
  stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                      ld_colours = c("white", "grey"))
@

Activity of dead or escaped animals is falsely scored as long series of zeros.
Our \texttt{sleepr} package offer a tool to detect and remove such artefactual data.
The updated version can be visualised in Fig~\ref{fig:fig-3}B.
<<>>=
# remove data after death
dt <- sleepr::curate_dead_animals(dt, moving)
# same as above
fig3B <- ggetho(dt, aes(y=label, z=moving)) +  
    stat_tile_etho() +                  
    stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
    stat_ld_annotations(x_limits = c(0, dt[,max(t)]), 
                        ld_colours = c("white", "grey"))
@

For the purpose of this example, we keep only individuals that have \emph{at least five days in LL}.
<<>>=
# for each id, we check for validity
valid_dt <- dt[ , .(valid = max(t) > days(5)), by=id]
# a vector of all valid ids
valid_ids <- valid_dt[valid == T, id]
# filter dt with the valid ids
dt <- dt[id %in% valid_ids]
summary(dt)
@

Note that as a result, we now have \Sexpr{nrow(dt[meta=T])} ``valid''individuals.

\subsection*{Double plotted actograms}
``Double-plotted actograms'' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In Fig~S1A, we show the double-plotted actograms of each animals:

<<>>=
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            stat_bar_tile_etho() + 
            facet_wrap( ~ label, ncol=4) +
            scale_y_discrete(name="Day")
@

\subsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
In several methods are implemented in \texttt{zeitbebr}. In this example, we generate $\chi{}^2$ periodrograms and lay them out in a grid (see \nameref{S1-Fig}).

<<>>=
dt_ll <- dt[t > days(1)]
per_dt <- periodogram(moving, 
                    	dt_ll, 
                    	resample_rate = 1/mins(10),
                    	FUN=chi_sq_periodogram)

per_dt <- find_peaks(per_dt)

figS1B <- ggperio(per_dt, aes(y = power, peak=peak)) + 
                  geom_line() +
                  geom_line(aes(y=signif_threshold), colour="red") + 
                  geom_peak() + 
                  facet_wrap( ~ label, ncol=4) 
@

<<r, eval=TRUE, echo=FALSE, message=FALSE, result="hide">>=
################################ CODE BELOW FOR COSMETIC ALTERATIONS OF FIGURES

library(cowplot)
my_theme <- theme(axis.title.y=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.line = element_blank(),
            panel.border = element_blank(),
            legend.position="none")

legend <- fig3A + theme(axis.title=element_blank(),
                        axis.text.x=element_blank(),
                        axis.ticks=element_blank(),
                        axis.line = element_blank(),
                        #axis.ticks.y=element_blank(),
                        panel.border = element_blank(),
                        #legend.text = "moving",
                        legend.position="left") + 
                  coord_cartesian(xlim=-days(c(10, 9))) + scale_y_discrete(position = "right")

fig3 <- cowplot::plot_grid(
                          legend,
                          fig3A +  my_theme,
                          fig3B +  my_theme,
                          nrow=1,
                          labels=c("", "A", "B"),
                          rel_widths = c(1, 2, 2),
                          label_size = 20,
                          axis="b",
                          align="h"
                          )
pdf("../fig/fig-3.pdf",w=12,h=12)
print(fig3)
dev.off()

figS1A <- figS1A +  theme(strip.text.x = element_text(size =8))
figS1B <- figS1B +  theme(strip.text.x = element_text(size =8))

figS1 <- cowplot::plot_grid(
                          figS1A,
                          figS1B,
                          nrow=1,
                          labels=c( "A", "B"),
                          label_size = 20,
                          align="v"
                          )

pdf("../fig/fig-S1.pdf",w=12,h=16)
print(figS1)
dev.off()
@