### 2018-04-04 manuscript.Rnw
fog 6 and a bit off proofreading

% Template for PLoS
% Version 3.4 January 2017
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

\usepackage{amsmath,amssymb}
\usepackage{changepage}
\usepackage[utf8x]{inputenc}
\usepackage{textcomp,marvosym}
\usepackage{cite}
\usepackage{nameref,hyperref}
\usepackage[right]{lineno}
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
\usepackage[table]{xcolor}
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

\reversemarginpar
\usepackage[colorinlistoftodos]{todonotes}
%\usepackage[colorinlistoftodos,disable]{todonotes}

%% Include all macros below
\newcommand{\citationneeded}[2][]{\todo[color=blue, fancyline, #1]{\textbf{Citation Needed:} #2}}
\newcommand{\TODO}[2][]{\todo[color=red, fancyline, #1]{\textbf{TODO:} #2}}

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Rethomics: an R framework to analyse high-throughput behavioural data} 
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Quentin Geissmann\textsuperscript{1*},
Luis G Rodriguez\textsuperscript{2},
Esteban J Beckwith\textsuperscript{1},
Giorgio F Gilestro\textsuperscript{1*}
\\
\bigskip
\textbf{1} Department of Life Sciences, Imperial College London, London, United Kingdom
\\
\textbf{2} Institute for Neuro- and Behavioral Biology, Westf{\"a}lische Wilhelms University, 48149 M{\"u}nster, Germany
\\
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* qgeissmann@gmail.com, giorgio@gilest.ro

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
%Ethomics, a quantitative and high-throughput approach to animal behaviour, is a new and exciting field.
The recent development of automatised methods that can score various behaviours on a large number of animals
provides biologists with an unprecedented set of tools to decipher these complex phenotypes. 
Analysing such data comes with several challenges that are largely shared across acquisition platform and paradigms.
%However, there is little effort in providing a generic toolbox to specifically analyse multiple and long behavioural time series.
%To address this limitation, we developed the \texttt{rethomics} framework, 
%We present the \texttt{rethomics} framework, 
%a suite of \texttt{R} packages that altogether offers utilities to work through all data manipulation stage:
%import, store, analyse and visualise behavioural data.
In the article herein, we present \texttt{rethomics}, a set of \texttt{R} packages that unifies analysis of behavioural dataset in an efficient and flexible manner.
%We implemented it in \texttt{R}\cite{r_core_team_r:_2017} since it is widely taught and adopted by computational biologists.
%It is also complemented with a vast ecosystem of open-source packages.
\texttt{rethomics} offers a computational solution to store, manipulate and visualise large amounts of behavioural data.
We propose it as a tool to fill the gap between behavioural biology and data sciences, thus connecting computational and behavioural scientists.
Our software comes with a extensive documentation as well as a set of both practical and theoretical tutorials (available \href{https://rethomics.github.io}{https://rethomics.github.io}).

%In this article, we describe it and show an example of its application to the study of circadian rhythms and blooming field of sleep fruit flies.
%The \texttt{rethomics} framework is available and documented at \href{https://rethomics.github.io}{https://rethomics.github.io}.

\listoftodos

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
%\section*{Author summary}

% TODO revert for sumbmisson
% \linenumbers
\TODO{revert line numbers before submission}
% TODO also embed bib

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
The behaviour of an animal is a complex phenotypical manifestation of the interaction between its nervous system, its internal state and its environment.
In the last few decades, our ability to record vast quantities of various phenotypical data has tremendously increased\cite{stephens_big_2015}.
The scoring of behaviours is certainly not an exception to this trend \cite{reiser_ethomics_2009}.
Indeed, many platforms have been developed in order to allow biologists to continuously record behaviours such as activity\cite{faville_how_2015}, position\cite{pelkowski_novel_2011} and feeding\cite{itskov_automated_2014,ro_flic:_2014} of single or multiple\cite{swierczek_high-throughput_2011,perez-escudero_idtracker:_2014} animals over long durations (days or weeks).

The availability of large amounts of data paves the way for in-depth quantitative analyses which, in turn, leads to the characterisation of new principles
and ultimately a better understanding of their underlying biology\cite{brown_study_2017,berman_measuring_2018}.
The multiplicity of model organisms, hypotheses and paradigms makes the existence of a diverse palette of recording tools important.
However, when it comes to the subsequent processing of the results, there is no unified, programmatic, framework that could be used as a set of building blocks in a pipeline.

Scripting interfaces are the standard in data sciences and statistics since they help delivering reproducible results \cite{peng_reproducible_2011,stodden_toward_2013}.
In addition, they can be used on remote resources such as computer clusters, which makes them more scalable to the context of `big data'\cite{hashem_rise_2015}.
For example, the field of bioinformatics benefits from sharing standard files formats, modular command line tools\cite{roumpeka_review_2017} and software packages\cite{huber_orchestrating_2015} that can be assembled to build pipelines\cite{leipzig_review_2017}.
Since many aspect of behaviour analysis are becoming increasingly linked to data sciences, the development of common tools and data structures would be valuable to the community.

At first, it may seem as though behavioural experiments are prohibitively heterogeneous -- in terms of model organisms, paradigm and time scale -- for an analogous ecosystem to arise.
However, some low-level conceptual consistencies and methodological challenges are common between experiments.
For instance, the results (\emph{i.e.} the `data')  feature a set of long time series (sometimes multivariate and irregular), but also contain a formal description of the treatment applied to each individual, the `metadata'.
Storing and accessing data and metadata efficiently involves the implementation of a nested data structure which, in principle,
can be shared between acquisition platforms and experimental paradigms.
% it broaden the scope for developing of utilities an shared principles.

In the article herein, we describe the \texttt{rethomics} platform, a collection of packages offering a solution to import, store, manipulate and visualise large amounts of behavioural results
in a data mining and statistical framework.
We also present two examples of application of this tool to the analysis of circadian rhythm in fruit flies, a widely studied behaviour.

\section*{Design and Implementation}
\texttt{rethomics} is implemented in \texttt{R}\cite{r_core_team_r:_2017}, which is widely taught and adopted by computational biologists,
as a collection of packages (Fig~\ref{fig:fig-1}).
This architecture follows the model of modern frameworks such as the \texttt{tidyverse}\cite{wickham_tidyverse:_2017}, which results in increased testability and maintainability.
In this model, each task of the analysis workflow (\emph{i.e.} data import, manipulation and visualisation) is handled by a different package.
At the core of \texttt{rethomics}, lies the \texttt{behavr} package, used to store large amounts data (\emph{e.g.} position and activity) and metadata (\emph{e.g.} treatment and genotype) in a unique \texttt{data.table}-derived object\cite{dowle_data.table:_2017}.
Any input package will import experimental data as a \texttt{behavr} table which can, in turn, be analysed and visualised regardless of the original input platform.
Visualisation and other results are standard objects, and can therefore be further analysed inside the wide \texttt{R} package ecosystem.

% Place figure captions after the first paragraph in which they are cited.

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-1.pdf}
	\caption{{\bf The \texttt{rethomics} workflow.}
		Diagram representing the interplay between, from left to right, the raw data, the \texttt{rethomics} packages (in blue) and the rest of the \texttt{R} ecosystem.}
	\label{fig:fig-1}
\end{figure}

\subsection*{Internal data structure}

Ethomics results can easily scale and data structure therefore gains from being computationally efficient – both in term of memory footprint and processing speed.
For instance, there could be very long time series, sampled several times per second, over multiple days, for each individual. %(typically, $k_i > 10^8, \forall i \in [1,n]$) 
In addition, time series can be multivariate, encoding coordinates, orientation, dimensions, activity, colour intensity and so on.
Furthermore, experiments may  feature a large number of individuals. %(typically $n > 100$).
Each individual is also associated with some metadata: a set of `metavariables' that describe experimental conditions.
For instance, metadata stores information regarding the date and location of the experiment, treatment, genotype, sex, \emph{post hoc} observations and other arbitrary metavariables.
A large set of metavariables is an important asset since they can later be used as covariates.

\texttt{behavr} tables link metadata and data within the same object, extending the syntax of \texttt{data.table} to manipulate, join and access metadata (Fig~\ref{fig:fig-2}A~and~B).
This approach guarantees that any data point can be mapped correctly to its parent metadata.
Furthermore, it allows implicit update of metadata when data is altered.
For instance, when is data filtered, only the remaining individuals should be in the new metadata. 
It is also important that metadata and data can interoperate.
For example, when one wants to update a variable according to the value of a metavariable (say, alter the variable $x$ only for animals with the metavariable $sex=`male'$).
The online tutorials and documentation provide a detailed set of examples and concrete use cases of \texttt{behavr}.

\begin{figure}[!h]
% 	\includegraphics[width=1\textwidth]{fig/fig-2.pdf}
	\caption{{\bf \texttt{behavr} table.}
	A: Illustration of a \texttt{behavr} object, the core data structure in \texttt{rethomics}.
		The metadata holds a single row for each of the $n$ individuals. 
		Its columns, the $p$ metavariables, are one of two kinds: either required -- and defined by the acquisition platform (\emph{i.e.} used to fetch the data) -- or user-defined (\emph{i.e.} arbitrary).
		In the data, each row is a `read' (\emph{i.e.} information about one individual at one time-point).
		It is formed of $q$ variables and is expected to have a very large number of reads, $k$, for each individual $i$.
		Data and metadata are implicitly joined on the \texttt{id} field.
		Note that the names used in this for variables and metavariable in this example are only plausible cases which will likely differ in practice. 
		B: Non exhaustive list of uses of a \texttt{behavr} table (referred as \texttt{dt}). 
		In addition to operations on data, which are inherited from \texttt{data.table},
		we provide utilities designed specifically to act on both metadata and data.  
		Commented examples are prefixed by `\texttt{>}'.
	}
	\label{fig:fig-2}
\end{figure}

\subsection*{Data import}
Data import packages translate results from a recording platform (\emph{e.g.} text files and databases) into a single \texttt{behavr} object.
Currently, we provide a package to read single or multi-beam Drosophila Activity Monitor System (Trikinetics Inc.) data and another one for Ethoscope data\cite{geissmann_ethoscopes:_2017}.
Although the structure of the raw results is very different, conceptually, loading data is very similar.
In all cases, the user is asked to generate a metadata table (one row per individual). 
In it, there will be both mandatory and optional columns.
The mandatory ones are the necessary and sufficient information to fetch data (\emph{e.g.} machine id, region of interest and date). 
The optional columns are user-defined arbitrary fields that translate experimental conditions (\emph{e.g.} treatment, genotype and sex).

In this respect, the metadata file is a standardised and comprehensive data frame describing an experiment.
It explicitly lists all treatments and individuals, which facilitates interspersion of conditions (indeed, without it, users are tempted to simplify their experimental design by, for instance, confounding device/location and treatment).
Furthermore, it streamlines the inclusion and analysis of further replicates in the same workflow.
Indeed, additional replicates can simply be added as new rows -- and the ID of the replicate later used, if needed, as a covariate.

\subsection*{Visualisation}

To integrate visualisation in \texttt{rethomics}, we implemented \texttt{ggetho},
a package that offers new tools that extent the widely adopted \texttt{ggplot2}\cite{wickham_ggplot2:_2016}.
\texttt{ggetho} makes full use of the internal \texttt{behavr} structure to summarise of temporal trends.
We implemented a set of new `layers' and `scales' that particularly applies to the visualisation of long experiments, with the ability to, for instance, display `double-plotted actograms', periodograms, annotate light and dark phases and wrap time over a given period. 
Importantly, \texttt{ggetho} is fully compatible with \texttt{ggplot2}. 
For instance, \texttt{ggplot2} operations such as faceting, transforming axes and adding new layers will function natively with \texttt{ggetho}.

<<'child-results.Rnw', child='results.Rnw'>>=
@

% We present an small example examining the circadian behaviour of 128 fruit flies in recorded in a DAM2 -- a paradigm very widely adopted. 
% A less formal description of this case and others are explained at \href{https://rethomics.github.io/}{https://rethomics.github.io/}.
% In order to prodive a more comprehensive and didactic example, data was altered and simplified.	
% Fig~\ref{fig:experiment} describes our case experiment (A) and the corresponding metadata (B).
% Briefly, three monitors were used each containing 16 males and 16 females of a different genotype (A, B and C).
% Two replicates were performed at different times.
% raw data and metadata files are publicly available TODO cite zenodo
%

\section*{Availability and Future Directions}
All packages in the \texttt{rethomics} framework are available under the terms of the GPLv3 license and listed at
\href{https://github.com/rethomics}{https://github.com/rethomics/}.
Extensive installation instructions as well as reproducible demos and tutorials are available at
\href{https://rethomics.github.io/}{https://rethomics.github.io/}.
All packages are continuously integrated and unit tested on several version of \texttt{R} to minimise the risk of present and future issues.

Several users, in different research groups, have already adopted and are contributing to the future development framework.
Several new packages in the \texttt{rethomics} framework are currently envisaged. 
They include utilities to input new behaviour tracking methods and analyse position and multi-animal interactions.

\section*{Supporting information}

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
\paragraph*{S1 Fig.}
\label{S1-Fig}
{\bf Complete version of Fig~\ref{fig:fig-4}.}
See Fig~\ref{fig:fig-4} for legend.

\section*{Acknowledgements}
We would like to thank people who have directly or indirectly contributed to the this manuscript.
In particular, Han Kim, for his invaluable comments on the early versions of \texttt{rethomics} and his dedicated contribution to the tutorials.
Maite Ogueta, for making the results data available.
\TODO{@esteban names of the BA users}
Alice French, Hannah Jones and Diana Bicazan for their comments as early users.
Marcus Gosh, for his feed back on the manuscript.
Brenna Williams, for her help to support multi-beam DAMS.
Patrick Kr{\"a}tschmer, for his time discussing the conceptual framework.
\TODO{@giorgio do you think it is ok to thank Rob?} Robert Dickinson, for his support.

% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
%\paragraph*{S1 Fig.}
%\label{S1_Fig}
%{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).

\nolinenumbers

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% or
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here. See http://journals.plos.org/plosone/s/latex for 
% step-by-step instructions.
%

\bibliography{manuscript}{}

%\begin{thebibliography}{10}
%\end{thebibliography}

\end{document}