### 2018-04-05 results.Rnw

<<echo=FALSE, message=FALSE>>=
library(knitr)
opts_knit$set(root.dir = "results")
setwd("results")
@

\section*{Results}
%In order to illustrate the workflow of \texttt{rethomics}, we provide a simplified and reproducible example, which could be readily modified to suit more intricate cases.

In order to illustrate the usefulness of \texttt{rethomics}, we provide two use cases examples.
The first one is a detailed and reproducible description of the loading and analysis activity data in the context of circadian rhythm, using DAM2 (Trikinetics Inc.) data.
The second one shows how \texttt{rethomics} integrates with the rest of \texttt{R} with the advanced case of a multi-scale analysis of a periodic behaviour, using continuous wavelet transform, on data generated with ethoscopes\cite{geissmann_ethoscopes:_2017}.

\subsection*{Drosophila Activity Monitor}
It uses the package \texttt{zeitgebr} which implements a suite of methods to compute autocorrelogram,
 $\chi{}^2$\cite{sokolove_chi_1978} and Lomb-Scargle\cite{ruf_lomb-scargle_1999} periodograms, and find their peaks.

The study of circadian rhythm, employing fruit flies as an animal model, is a well established research field recently awarded with the Nobel Prize in Physiology or Medicine.
Importantly, DAM2 (the second, and most popular, version of the Drosophila Activity Monitors System) is the most widely adopted behaviour recording platform in this research field.
We gathered a subset of the data from a recent publication\cite{buhl_quasimodo_2016}, kindly made publicly available by the authors\cite{ogueta_ll_2018}.
Wild type flies are highly rhythmic in Light-Dark (LD) cycles and become arrhythmic in constant light (LL).
In their study, the authors gain understanding of the function of the molecular clock by showing that overexpression of the gene NKCC makes the flies rhythmic in both LD and LL,
and that the endogenous period in LL is longer than 24 hours.

We took two genotypes employed in the study, one control group (NKCC\textsuperscript{ox}/+) and one where NKCC\textsuperscript{ox} is overexpressed in all clock neurons (TIM/NKCC\textsuperscript{ox}).
In particular, we re-analysed two repetitions of the same experiment in which a total of 58 animals were recorded for three to four days in LD and then subjected to constant light for six or seven days.
The \texttt{metadata.csv} file, which describes exhaustively all individuals, as well as all the associated result files can be downloaded at \href{https://zenodo.org/record/1172980}{https://zenodo.org/record/1172980}.
As a prerequisite, we downloaded the data and extracted the zip archive in our working directory.
We start the analysis by loading the necessary \texttt{rethomics} packages (see the webpage for installation instructions):

<<message=FALSE>>=
library(damr)      # input DAM data
library(zeitgebr)  # periodogram computation
library(sleepr)    # sleep analysis
library(ggetho)    # behaviour visualisation
@

Then, the metadata file is read and linked to the \texttt{.txt} result files.

<<cache=TRUE>>=
metadata <- link_dam_metadata("metadata.csv", ".")   # linking
# print(metadata)                                     # check metadata
dt <- load_dam(metadata)                             # loading
summary(dt)                                           # quick summary
@

\subsubsection*{Preprocessing}
The two replicates do not have the same baseline time and we would like to express the time relative to the important event:
the transition from LD to LL. 
Therefore, we subtract the \texttt{baseline\_days} metavariable from the \texttt{t} variable.
This gives us an opportunity to illustrate the use \texttt{xmv()}, which expands metavariables as variables.
In addition, we use the \texttt{data.table} syntax to create, in place, a \texttt{moving} variable.
It is \texttt{TRUE} when and only when \texttt{activity} is greater than zero:

<<message=F, results="hide", cache=TRUE>>=
# baseline subtraction -- note the use of xmv
dt[,t := t - days(xmv(baseline_days))]
dt[, moving :=  activity > 0]  
@

<<cache=TRUE>>=
summary(dt)
@

To simplify visualisation, we create our own \texttt{label} metavariable, as the combination of a number and \texttt{genotype}.
In the restricted context of this analysis, \texttt{label} acts as a unique identifier.
Importantly, we also keep \texttt{id} as an \emph{unambiguous} unique identifier.
Indeed, two animals in separate experiments may have the same label, but different \texttt{id}s.
in addition, if the metadata changes -- for instance by the addition or removal of individuals -- the label is likely to change, not the \texttt{id}.
%,which could lead to confusion.

<<message=FALSE, results="hide", cache=TRUE>>=
dt[, label := interaction(1:.N, genotype), meta = T]
print(dt)
@

\subsubsection*{Curation}
It is important to visualise an overview of how each individual behaved and, if necessary, alter the data accordingly. For this, we generate a tile plot (Fig~\ref{fig:fig-3}A).

\begin{figure}[!h]
%    \includegraphics[width=1\textwidth]{fig/fig-3.pdf}
	\caption{{\bf Experiment quality control.}
			Tile plot showing the fraction of time spent moving as a colour intensity.
			Each individual is represented by a row and time, on the x-axis, is binned in 30 minutes.
			A: Uncurated raw data.
			B: Data after the curation step. Time was trimmed and data from dead animals removed. 
			Red `\textbf{+}' symbols show animals that were removed from the subsequent analysis as they had less than five complete days in LL.}
	\label{fig:fig-3}
\end{figure}

<<cache=TRUE>>=
# make a ggplot object with label on the y and moving on the z axis
fig3A <- ggetho(dt, aes(y = label, z = moving)) +  
  # show data as a tile plot
  # that is z is a pixel whose intensity maps moving
  stat_tile_etho() +                  
  # add layers to draw annotations to show L and D phases
  # as white and black, respectively
  # the first layer is for the baseline (until t = 0)
  stat_ld_annotations(x_limits = c(dt[,min(t)], 0)) +
  # in the 2nd one, we start at 0 and use grey 
  # instead of black as we work in LL
  stat_ld_annotations(x_limits = c(0, dt[, max(t)]), 
                      ld_colours = c("white", "grey"))
@

The activity of dead or escaped animals is falsely scored as long series of zeros (see, for instance, individual 30 and 18 in Fig~\ref{fig:fig-3}A).
The \texttt{sleepr} package offers a tool to detect and remove such artefactual data.

<<cache=TRUE>>=
# remove data after death
dt <- curate_dead_animals(dt, moving)
@

In addition, we can trim our data to have the same number of days across experiments and individuals.
<<cache=TRUE>>=
# filter dt between -2d and 6d
dt <- dt[t %between% days(c(-2, 6))]
# same as above
fig3B <- ggetho(dt, aes(y = label, z = moving)) +  
    stat_tile_etho() +                  
    stat_ld_annotations(x_limits = c(dt[, min(t)], 0)) +
    stat_ld_annotations(x_limits = c(0, dt[, max(t)]), 
                        ld_colours = c("white", "grey"))
@

For the purpose of this example, we also exclude animals that died prematurely, and keep only individuals that have \emph{at least five days in LL}. An overview of the curate data can be visualised in Fig~\ref{fig:fig-3}B.

<<cache=TRUE, echo=FALSE>>=
# draw point for invalid animals
valid_dt <- dt[, .(valid = max(t) > days(5)), by = id]
valid_ids <- valid_dt[valid == T, id]
fig3B <- fig3B + geom_text(data = dt[!id %in% valid_ids,.(label), meta = T], aes(x = days(6), y = label), inherit.aes = F, colour = "red", label = "+", fontface = "bold")
@

<<cache=TRUE>>=
# for each id, we check for validity
valid_dt <- dt[, .(valid = max(t) > days(5)), by = id]
# a vector of all valid ids
valid_ids <- valid_dt[valid == T, id]
# filter dt with the valid ids
dt <- dt[id %in% valid_ids]
summary(dt)
@

Note that as a result, we now have \Sexpr{nrow(dt[meta=T])} `valid' individuals.

\subsubsection*{Double-plotted actograms}
`Double-plotted actograms' are a common visualisation of periodicity and rhythmicity in circadian experiments.
In  \nameref{S1-Fig}A, we show a double-plotted actogram for each animal.
A selected sample of four individuals for each genotype is shown in Fig~\ref{fig:fig-4}A.

<<cache=TRUE, warning=FALSE, result="hide", message=FALSE>>=
# we also show a subset of this figure in 4A
figS1A <- ggetho(dt, aes(z = moving), multiplot = 2) +
            # bars n the z axis could 
            # one could also use stat_tile_etho
            stat_bar_tile_etho() + 
            # split plot by individual
            facet_wrap( ~ label, ncol = 4) +
            # rename the y axis
            scale_y_discrete(name = "Day")
@

<<cache=TRUE, echo=FALSE, warning=FALSE, message=FALSE, result="hide">>=
#fig_4_picked_labels = paste(c(5,10,33,39, 13,17,22,51), rep(c("NKCCOX/+", "Tim/NKCCOX"), each=4), sep=".")
helper_dt_picked_labs <- data.table(num = c(5,10,33,39, 13,17,22,51),
                                    genotype =  rep(c("NKCCOX/+", "Tim/NKCCOX"), each=4))
helper_dt_picked_labs[,label := interaction(num, genotype)]

  # add a bg that depends on labels
bg_layer =   geom_rect(data=helper_dt_picked_labs, aes(fill=genotype),
            xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf,
            alpha = 0.3, inherit.aes = F)

fig_4_picked_labels = helper_dt_picked_labs[,label]

fig4A <- ggetho(dt[xmv(label) %in% fig_4_picked_labels],
						 aes(z = moving), multiplot = 2) +
  bg_layer+
	stat_bar_tile_etho() + 
	facet_wrap( ~ label, ncol = 4) +
	scale_y_discrete(name = "Day") +
  theme(strip.text = element_text(size = 11))

ylabs <- ggplot_build(fig4A)$layout$panel_ranges[[1]]$y.labels

new_labs <- ifelse(1:length(ylabs) %%2, ylabs, "")

fig4A <- fig4A + scale_y_discrete(name = "Day",
                                  labels = new_labs
                                  ) + stat_ld_annotations(ld_colours = c("white", "grey"))
@

\subsubsection*{Periodograms}
Ultimately, in order to quantify periodicity and rhythmicity, we compute periodograms.
Several methods are implemented in \texttt{zeitbebr}: $\chi{}^2$, Lomb-Scargle, autocorrelation 
and Fourier.
In this example, we generate $\chi{}^2$ periodograms and lay them out in a grid.
Periodograms for the subset of eight animals used in Fig~\ref{fig:fig-4}A is shown in Fig~\ref{fig:fig-4}B. 
See \nameref{S1-Fig}B for the visualisation of all individuals.

<<cache=TRUE>>=
# only the LL data
dt_ll <- dt[t > days(1)]
# compute chi sqr periodogram 
per_dt <- periodogram(moving, 
                    	dt_ll, 
                    	resample_rate = 1 / mins(10),
                    	FUN=chi_sq_periodogram)

per_dt <- find_peaks(per_dt)
# we also show a subset of this figure in supplementary materials
figS1B <- ggperio(per_dt, aes(y = power, peak = peak)) + 
                  # periododogram drawn as a line
                  geom_line() +
                  # the significance line in red
                  geom_line(aes(y = signif_threshold), colour = "red") + 
                  # point and text at the peak
                  geom_peak() + 
                  # divide plot by individual
                  facet_wrap( ~ label, ncol = 4) 
@

<<cache=TRUE, echo=FALSE, warning=FALSE>>=
fig4B <- ggperio(per_dt[xmv(label) %in% fig_4_picked_labels],
                      aes(y = power, peak = peak)) + 
            # periododogram drawn as a line
            geom_line() +
            bg_layer +
            # the significance line in red
            geom_line(aes(y = signif_threshold), colour = "red") + 
            # point and text at the peak
            geom_peak() + 
            # divide plot by individual
            facet_wrap( ~ label, ncol = 4) + 
            scale_y_continuous(name = "Power") +
            theme(strip.text = element_text(size = 11))

@

\begin{figure}[!h]
%	\includegraphics[width=1\textwidth]{fig/fig-4.pdf}
	\caption{{\bf Visualisation of the periodicity in activity of eight selected animals.}
		A: Double-plotted actograms showing activity over the experiment. Transition from LD to LL happens at day 0.
		B: $\chi{}^2$ periodograms over the LL part of the experiment matching the animals in A.
		The blue cross represents the first peak (if present) above the significance threshold (red line).
		Titles on top of each facet refer to the label allocated to each individual.
		See \nameref{S1-Fig} for all \Sexpr{nrow(dt[meta=TRUE])} animals.
	}
	\label{fig:fig-4}
\end{figure}

\subsubsection*{Population statistics}

As shown in the original study, double-plotted actograms and periodograms suggest that NKCC\textsuperscript{ox}/+ flies are mostly 
arhythmic in LL whilst Tim/NKCC\textsuperscript{ox} appear to have a consistent, long-period rhythm.
To visualise this at the population scale, we can plot an average periodogram (see Fig~\ref{fig:fig-5}A):

<<cache=TRUE, warning=FALSE,message=FALSE, result="hide">>=
# display periodogram 
fig5A <- ggperio(per_dt, aes(y = power - signif_threshold,
                             colour = genotype)) + 
          # periododogram shown as a line for population mean
          # and bootstrap error bars
          stat_pop_etho(method = ggplot2::mean_cl_boot) +
          # rename x and y axis 
          scale_y_continuous(name = "Relative power") + 
          scale_x_hours("Period")
@

\begin{figure}[!h]
%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Population statistics on circadian phenotype.}
			A: Average periodograms. 
			      The aggregated relative power of the periodogram of all animals.
			      The solid lines and the shaded areas show population means and their 95\% bootstrap confidence interval, respectively.
			B: Frequencies of rhythmic animals.
			      Number of rhythmic animals (\emph{i.e.} with a significant peak) in each genotypes.
			      Dark and clear fillings indicate rhythmic and arhythmic animals, respectively.
      C: Peak periodicity power and average.
			      Values of the peak period for animals with a significant peak (\emph{i.e.} rhythmic).
			      Individual animals are shown by dots whose size represent relative power of the peak period.
			      The error bars are 95\% bootstrap confidence interval on the population mean.
			      }
	\label{fig:fig-5}
\end{figure}

To further quantify this difference, we can show the number of rhythmic animals -- \emph{i.e.} individuals for which a peak was found -- in each group (see Fig~\ref{fig:fig-5}B).
Then, we can compare the average value of the peak for the rhythmic animals (see Fig~\ref{fig:fig-5}C).
First of all, we compute a summary per individual (\texttt{by=id}):

<<cache=TRUE, warning=FALSE, results="hide">>=
summary_dt <- 
    per_dt[, 
          .(
           first_peak_period = period[peak == 1],
            # {} can be used for tmp variables
           first_peak_rel_power = {
             signif = signif_threshold[peak == 1]
             power = power[peak == 1]
             power - signif
            },
           is_rhythmic = any(peak == 1)
           ),
           by=id]

# rejoin metadata
summary_dt <- rejoin(summary_dt)
@

\texttt{summary\_dt} is just a regular data frame with one row per individual, containing both metadata and our summary statistics. It can therefore be used directly by \texttt{ggplot} and other tools:

<<cache=TRUE, warning=FALSE, results="hide">>=
# standard ggplot
fig5B <- ggplot(summary_dt, aes(x = genotype,
                                fill = genotype,
                                alpha = is_rhythmic
                                )) +
              geom_bar(colour="black")

@

<<>>=
# standard ggplot
fig5C <- ggplot(summary_dt, aes(y = first_peak_period,
                                x = genotype)) + 
              # draw the mean of each genotype group
              stat_summary(fun.y = mean, geom = "point", shape=3) +
              # draw bootstrap confidence intervals
              stat_summary(fun.data = mean_cl_boot, geom = "errorbar") +
              # shows all individuals as points
              # the size of the point expresses the power of the peak
              geom_jitter(aes(colour = genotype, 
                              size = first_peak_rel_power),
                              alpha = 0.67) +
              # We would like to convert time in hour
              scale_y_hours("Period")
@

\texttt{R} provides one of the richest statistical toolbox available, which allows users to go deeper in the analysis of the extracted variables.
One could, for instance, perform a $\chi{}^2$ test on the number of rhythmic \emph{vs} arhythmic flies in both genotypes.
To address the same question, we fit a binomial generalised linear model:

<<>>=
fit <- glm(is_rhythmic ~ genotype, summary_dt, family = "binomial")

summary(fit)$coefficients
@

<<echo=FALSE>>=
str <- sapply(strsplit(format(summary(fit)$coefficients[2,4], scientific=TRUE, digits = 3), split="e"), function(x) paste0("$", x[1], " \\times 10^{", x[2], "}$"))
@

The result shows a strong positive effect of genotype Tim/NKCC\textsuperscript{ox} on the probability of being rhythmic ($p$-value \Sexpr{str}):

Lastly, we can generate a table that compute arbitrary population statistics for each genotype:
<<warning=FALSE>>=
result_dt <- summary_dt[,
          .(
            mean_period = round(mean(first_peak_period, na.rm = T) / hours(1), 2),
            sd_period = round(sd(first_peak_period, na.rm = T) / hours(1), 2),
            percent_rhythmic = round(100 * sum(is_rhythmic) / .N, 2),
            n_rhythmic = sum(is_rhythmic),
            n = .N
            ),
          by = genotype
          ]
result_dt
@

This example shows how \texttt{rethomics} can be used from the raw data to making publication-quality figures and statistics.
We were able to comprehensively analyse the data from a circadian experiment with a few lines of code.
This workflow applies particularly to much larger data sets and provides a large degree of flexibility that will allow biologists to tune their analysis to their specific questions.

\subsection*{Multi-scale analysis of position}

One of the challenges of behaviour analysis is the `nesting' of events happening over different time scales.
In other words, a behavioural variable can be modulated by the circadian rhythm, but also by co-occurring ultradian and infradian rhythms.
For instance, an animal could have rhythmic bursts of locomotor activity recurring at high frequency (\emph{e.g.} 1~min), but also a circadian  (\emph{i.e.} 24~h) regulation of the same variable.
In this example, both rhythms happen at time scales separated by approximately three orders of magnitudes, which makes them difficult to visualise and integrate in the same analysis.
Being able to keep frequency information over multiple scales is however important in some cases.
In particular, when interested in the frequency modulation of one rhythm by another --
that is, if the periodicity of a high frequency rhythm itself can be function of a lower frequency one.
%For instance, the frequency of rhythmic activity bursts could itself be regulated by another rhythm such as the circadian rhythm.

The problem of understanding time series at different scales in not uncommon in fields such as economics\cite{aguiar-conraria_business_2011}, climate sciences\cite{lau_climate_1995} and ecology\cite{cazelles_wavelet_2008} where variables are governed by multiple underlying rhythms (\emph{e.g.} tidal, daily, yearly and multi-yearly).
One approach is to study a variable of interest in the time/period domain using, for instance, continuous wavelet transform (CWT)\cite{grossmann_decomposition_1984}.
In the context of chronobiology, CWT has been suggested as a tool to investigate ultradian rhythms \cite{leise_wavelet_2013}.

To illustrate how \texttt{rethomic} integrates with other packages to render type of less traditional analysis possible,
we performed a wavelet analysis of the position of 80 naive fruit flies (40 females and 40 males) in their glass tubes.
We used the package \texttt{scopr}, part of \texttt{rethomics}, to load five days of ethoscope positional data, which we sampled at 0.1 Hz.
Our variable of interest is the position of animals in their tube (from the food end, 0,  to the cotton end, 1).
Fig~\ref{fig:fig-6}A-C shows the raw position data at two different scales for two representative animals.

\begin{figure}[!h]
	%  	\includegraphics[width=1\textwidth]{fig/fig-5.pdf}
	\caption{{\bf Wavelet analysis of positional data.}
		A: Raw position data for a representative female (top) and male (bottom) \emph{Drosophila} over five days, in black. 
		The thick coloured lines show the average position every two.
		The green rectangles in the background shows the two time windows selected for B and C.
		B: Close up of A, showing position over one hour, in the beginning of the L phase of day 1.
		C: Close up of A, showing position over one hour, in the middle of the L phase of day 1.
		D: Continuous wavelet transform spectrogram for the two representative animals.
		E: Average spectrogram across 40 males and 40 females.
		In D and E, the lines on the right shows the marginal power spectra corresponding to the shown spectrograms (average across all time).
		The male data was collected and described in our previous study \cite{geissmann_ethoscopes:_2017} (controls in figure~5M-P) and the females data was acquired in parallel, in the same experimental conditions, but not previously published.
	}
	\label{fig:fig-6}
\end{figure}

In order to compute CWT, we used the \texttt{WaveletComp} package\cite{schmidbauer_waveletcomp:_2018}.
We then averaged the result of the five consecutive days in the time/period domain over one circadian day both for the two representative animals (Fig~\ref{fig:fig-6}D)
and for the population(Fig~\ref{fig:fig-6}E).

As suggested by the slow oscilations of the mean position (Fig~\ref{fig:fig-6}A), we observe strong, high period (12~h and 24~h), rhythms.
In addition, a large amount of signal is detected for low period (around 60~s) events.
This correspond to the position of animals walking along (back and forward) their tubes in a very paced manner.

Interestingly, in females, this low period pace appears to be frequency modulated during the L phase, suggesting a slower walking speed around ZT6~h.
In contrast, males shows only a high frequency rhythm around the phase transitions (L to D and D to L).
Surprisingly, the peak of high frequency rhythm imply a faster pace in males (approximately 60~s) than in females (approximately 120~s) --
indicating that, when active, males walk faster than females.

This non-exhaustive proof of principle illustrates how analysis of behavioural data can be taken further by adapting the wide range of numerical tools already available in the \texttt{R} ecosystem.

<<r, eval=TRUE, echo=FALSE, message=FALSE, result="hide",cache=TRUE, warning=FALSE>>=
################################ CODE BELOW FOR COSMETIC ALTERATIONS OF FIGURES
library(cowplot)
label_size <- 24 # "A", "B" ...
my_theme <- theme(axis.title.y=element_blank(),
            axis.text.y=element_blank(),
            axis.ticks=element_blank(),
            axis.line = element_blank(),
            panel.border = element_blank(),
            legend.position="none")

legend <- fig3A + theme(axis.title=element_blank(),
                        axis.text.x=element_blank(),
                        axis.ticks=element_blank(),
                        axis.line = element_blank(),
                        axis.text.y = element_text(size=8),
                        #axis.ticks.y=element_blank(),
                        panel.border = element_blank(),
                        #legend.text = "moving",
                        legend.position="none") + 
                  coord_cartesian(xlim=-days(c(10, 9))) + scale_y_discrete(position = "right")

fig3 <- cowplot::plot_grid(
                          legend,
                          fig3A +  my_theme,
                          fig3B +  my_theme,
                          nrow=1,
                          labels=c("A", "", "B"),
                          rel_widths = c(1, 4, 4),
                          label_size = label_size,
                          axis="b",
                          align="h",
                          hjust=0, vjust=1.01
                          )
fill_legend <- 
  cowplot::plot_grid(NULL,
    cowplot::get_legend(
        fig3A + theme(legend.position = "bottom", legend.direction = "horizontal") + scale_fill_continuous(labels = scales::percent, name="Moving ", breaks=c(0,.5,1))
        ), NULL
    )
#?theme

fig3_final <- cowplot::plot_grid(fig3, fill_legend,nrow = 2, rel_heights = c(15,1))

pdf("../fig/fig-3.pdf",w=12,h=8)
print(fig3_final)
tmp  <- dev.off()

#=================================================
figS1A <- figS1A +  theme(strip.text.x = element_text(size =8))
figS1B <- figS1B +  theme(strip.text.x = element_text(size =8))

figS1 <- cowplot::plot_grid(
                          figS1A,
                          figS1B,
                          nrow=1,
                          labels=c( "A", "B"),
                          label_size = label_size,
                          align="v",
                          axis="b",
                          hjust=0, vjust=1.01
                          )

pdf("../fig/fig-S1.pdf",w=12,h=16)
print(figS1)
tmp <- dev.off()

#===================================================

fig4 <- cowplot::plot_grid(
				fig4A + theme(legend.position = "none"),
				fig4B+ theme(legend.position = "none"),
				nrow=1,
				labels=c("A", "B"),
				align="v",
				axis="b",
				label_size = label_size,
				hjust=0, vjust=1.01
)

fill_legend <- 
  cowplot::plot_grid(NULL,
    cowplot::get_legend(
        fig4A + theme(legend.position = "bottom", legend.direction = "horizontal") + scale_fill_discrete(name="Genotype "))
        )
#?theme

fig4_final <- cowplot::plot_grid(fig4, fill_legend,nrow = 2, rel_heights = c(15,1))
pdf("../fig/fig-4.pdf",w=12,h=4)
print(fig4_final)
tmp  <- dev.off()

#=================================================

fig5 <- cowplot::plot_grid(fig5A + theme(legend.position="none"),
                           fig5B + theme(legend.position="none") + scale_y_continuous(name="Number") + scale_x_discrete(name="Genotype"),
                           fig5C + theme(legend.position="none") + scale_x_discrete(name="Genotype"),
                           nrow=1,
                           label_size = label_size,
                           align = "hv",
                           axis="b",
                           labels=c( "A", "B", "C"),
                           hjust=0, vjust=1.01
                           #?cowplot::plot_grid
                           )

dummy <- ggplot(summary_dt, aes(is_rhythmic, first_peak_period, size= first_peak_power, fill= genotype, colour=genotype)) + 
    geom_bar(aes(x=genotype, alpha= ifelse(is_rhythmic, "Rhythmic", "Arhythmic"), fill=genotype), inherit.aes = F) + 
  geom_jitter(aes(is_rhythmic, first_peak_period, size=first_peak_rel_power), inherit.aes = F) +
   scale_fill_discrete(name="Genotype")+scale_size(name= "Peak relative power") +  scale_alpha_discrete(name= "") +
  theme(legend.position = "bottom")  + 
  guides(fill = guide_legend(order=1),
         alpha = guide_legend(order=2),
         size = guide_legend(order=3))

legend <- plot_grid(NULL, get_legend(dummy), NULL, nrow=1)

fig5_final <- cowplot::plot_grid(fig5, legend, nrow = 2, rel_heights = c(15,1))
pdf("../fig/fig-5.pdf",w=12,h=4)
print(fig5_final)
tmp <- dev.off()
@